An Elastica-driven Digital Curve Evolution Model for Image Segmentation\thanks{This work has been partially funded by CoMeDiC ANR-15-CE40-0006 research grant.}

### 2019-07-08 jmiv.tex
fatigue

\HT{The use of curvature for surface regularization in the case of the Willmore energy was studied
  in~\cite{bobenko2005discrete}, although results look promising, authors used an inexact linearized version of the
  square curvature at every step.}

@@

In image segmentation, Zehiry and Grady have shown that injecting curvature as a regularizer can help recover thin
and elongated objects \cite{zehiry10fast}. Similarly Schoenemann {\em et al.}  \cite{schoenemann09linear} have
considered the elastica energy in their ratio-based image segmentation approach. In~\cite{schoenemann2011elastic},
Schoenemann~{\em et al.} extend ratio-cut approaches to segmentation with a curvature term. Their global optimization

framework shows the considerable \HT{advantages} of using such regularizer in common binary image segmentation tasks.
However the time complexity of their algorithm \HT{is prohibitive,} even if curvature is \HT{again} coarsely
approximated.  \HT{In more recent work, Lim {\em et al.} propose an edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set formulation. They observe improved results compared
to~\cite{casseles97geodesic} and other perimeter-based priors, but observe that a shape prior and a close-enough
solution must also be provided to achieve good results.}

@@

Segmentation methods using Elastica or Willmore energies \HT{can be formulated via the computation of} corresponding
flow, i.e. following its gradient. Well-founded level-set methods involving Willmore energy \cite{droske2004level} have
been used in medical image segmentation \cite{lim2012introducing}. Another level-set formulation similar to the
Chan-Vese model \HT{was} proposed in \cite{zhu2013image}. Willmore flow can also be carried out by phase-field models
(see \cite{bretin2015phase}) but they are less suited to image segmentation since interfaces are blurred in such
models. Threshold dynamics can also be considered for such energies \cite{esedoglu2008threshold} and has been proposed
for image desocclusion \cite{esedoglu2005threshold}. \HT{However, flow-based methods will always deliver
  a local minimum, which heavily depend on initialization for good results}.

@@

In this paper we are interested in purely discrete variational segmentation models involving an elastica
energy. Recently, new estimators of curvature along digital contours have been proposed
\cite{roussillon11mdca,coeurjolly13integral,schindele17mdca} with the desirable multigrid convergence property. This
motivates us to propose models in which they can be used successfully.

\HT{In the remainder, we} investigate the use of the digital integral invariant curvature estimator
\cite{coeurjolly13integral} in a discrete variational segmentation framework. More precisely, we show how to incorporate
it in a digital flow minimizing its squared curvature. This model is expressible as a discrete combinatorial
optimization model, which we solve using a classical variant of QPBO algorithm \cite{rother07qpbo}. \HT{Our solution
  does not} compute a global solution to the whole domain, but our approach leads to a digital flow similar to
continuous \HT{Elastica}, as shown by our experiments. We present an application of the model as a post-processing step
in a segmentation framework, and demonstrate how it can improve standard results. 
\daniel{This paper is an extended version of a previous work that appeared in \cite{antunes19}. We've included a deeper discussion on the optimization method we've used and we present a comparison of our segmentation method with two other related works, and illustrate the results with several examples. Moreover, we have included a new local combinatorial optimization approach for elastica minimization. Finnaly,  the code produced in this paper is freely available on github\footnote{https://www.github.com/danoan/BTools}.}

\begin{itemize}
%% \item{Geometric properties and regularization}
%% \item{The role of curvature}

\item{What are the difficulties in use curvature as a regularizer (point out convex relaxation approaches) and how a digital approach may help us }
\item{Outline: We present two models that illustrate how digital estimators can be used to minimize the elastica energy on digital sets.  }
\item{\todo{Add related works : geometric priors, binary segmentation, other digital curve evolution models}}
\item{\todo{Talk about Willmore flow with phase field or with threshold dynamics}}

\item{\textbf{Constraint optimization}}
  \begin{itemize}
 %% \item{Curvature regularity for region-based image segmentation and inpainting: A linear programming relaxation \cite{schoenemann09linear} }	
  \item{A Fast Algorithm for Euler's Elastica Model Using Augmented Lagrangian Method \cite{tai11elastica} }		
  %%\item{The elastic ratio: Introducing curvature into ratio-based image segmentation \cite{schoenemann2011elastic}
  %%}		%% DONE
  \item{Curvature Regularization for Curves and Surfaces in a Global Optimization Framework \cite{strandmark11globalframework} }
  \end{itemize}

\item{\textbf{Discrete approaches}}
  \begin{itemize}
  \item{Fast global optimization of curvature \cite{zehiry10fast}	}
  \item{Efficient Squared Curvature \cite{nieuwenhuis14efficient} 	}
  \item{Contrast driven {E}lastica for image segmentation \cite{el2016contrast} }	
  \end{itemize}

\end{itemize}

\textit{Outline}. Section two reviews the concept of multigrid
convergence and highlights its importance for the definition of
digital estimators. Next, we describe two convergent estimators used
in this paper, one for tangent and the other for curvature. They are
used in the optimization model and in the definition of the digital
elastica. Section three describes a local combinatorial optimization model
suitable for different curvature estimators. Section four describes the proposed 
curvature evolution model along with several illustrations of digital flows. Section five
explains how to use the evolution model as a post-processing step in
an image segmentation framework.

@@

>>>>

\subsection{Existing works}
\HT{The use of curvature for surface regularization in the case of the Willmore energy was studied
  in~\cite{bobenko2005discrete}, although results look promising, authors used an inexact linearized version of the
  square curvature at every step.}

@@

In image segmentation, Zehiry and Grady have shown that injecting curvature as a regularizer can help recover thin
and elongated objects \cite{zehiry10fast}. Similarly Schoenemann {\em et al.}  \cite{schoenemann09linear} have
considered the elastica energy in their ratio-based image segmentation approach. In~\cite{schoenemann2011elastic},
Schoenemann~{\em et al.} extend ratio-cut approaches to segmentation with a curvature term. Their global optimization
framework shows the considerable \HT{advantages} of using such regularizer in common binary image segmentation tasks.
However the time complexity of their algorithm \HT{is prohibitive,} even if curvature is \HT{again} coarsely
approximated.  \HT{In more recent work, Lim {\em et al.} propose an edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set formulation. They observe improved results compared
to~\cite{casseles97geodesic} and other perimeter-based priors, but observe that a shape prior and a close-enough
solution must also be provided to achieve good results.}

@@

Segmentation methods using Elastica or Willmore energies \HT{can be formulated via the computation of} corresponding
flow, i.e. following its gradient. Well-founded level-set methods involving Willmore energy \cite{droske2004level} have
been used in medical image segmentation \cite{lim2012introducing}. Another level-set formulation similar to the
Chan-Vese model \HT{was} proposed in \cite{zhu2013image}. Willmore flow can also be carried out by phase-field models
(see \cite{bretin2015phase}) but they are less suited to image segmentation since interfaces are blurred in such
models. Threshold dynamics can also be considered for such energies \cite{esedoglu2008threshold} and has been proposed
for image desocclusion \cite{esedoglu2005threshold}. \HT{However, flow-based methods will always deliver a local
  minimum, which heavily depend on initialization for good results}.  \HT{A similar approach is to formulate the
  Elastica as a regularization in a variational framework and use operator splitting to optimize it. Tai {\em et
    al.}~\cite{tai11elastica} and more recently Deng~{\em et al.}~\cite{deng2019new} have proposed this approach, which
  is well-suited to inverse problem solving. The problem remains non-convex and the nature of the converged iterate
are not studied, i.e. how close they are to a global optimum.}

@@

\HT{Discrete approaches have been used to tackle the Elastica difficulties. As noted before~\cite{schoenemann09linear}
  have linearized a mesh-based approach, which was expanded upon by~\cite{strandmark11globalframework}, while
  \cite{zehiry10fast} have used a binary solver for a non-submodular approach on a regular square grid of pixels, which
  was expanded in~\cite{el2016contrast}. In~\cite{nieuwenhuis14efficient} , a specific solver is developed to yield a
  better solution to an extension of the formulation in~\cite{zehiry10fast}.}

\subsection{Motivation}
\HT{In spite of the multiplicity of existing approaches, it seems none is currently completely satisfactory. While the
  sought solution is expected to be very smooth based on geometrical expectations, the mathematical properties
  of the related operators make this smoothness elusive. Non-convexity and the presence of high-order terms make
  Elastica difficult to optimize and to approximate, let alone compute efficiently. The computed solution in the
  literature are rarely evaluated, and the complexity of the proposed solution is often high.}

%%\item{What are the difficulties in use curvature as a regularizer (point out convex relaxation approaches) and how a digital approach may help us }
%%\item{Outline: We present two models that illustrate how digital estimators can be used to minimize the elastica energy on digital sets.  }
%%\item{\todo{Add related works : geometric priors, binary segmentation, other digital curve evolution models}}
%%\item{\todo{Talk about Willmore flow with phase field or with threshold dynamics}}

\HT{Consequently,} we are interested in \HT{studying} purely discrete variational segmentation models involving an
elastica energy. Recently, new estimators of curvature along digital contours have been proposed
\cite{roussillon11mdca,coeurjolly13integral,schindele17mdca} with the desirable multigrid convergence property. This
motivates us to propose models in which they can be used successfully.

\HT{In the remainder, we propose to} investigate the use of the digital integral invariant curvature estimator
\cite{coeurjolly13integral} in a discrete variational segmentation framework. More precisely, we show how to incorporate
it in a digital flow minimizing its squared curvature. This model is expressible as a discrete combinatorial
optimization model, which we solve using a classical variant of QPBO algorithm \cite{rother07qpbo}. \HT{Our solution
  does not} compute a global solution to the whole domain, but our approach leads to a digital flow similar to
continuous \HT{Elastica}, as shown by our experiments. We present an application of the model as a post-processing step
in a segmentation framework, and demonstrate how it can improve standard results. 
\daniel{This paper is an extended version of a previous work that appeared in \cite{antunes19}. We've included a deeper
  discussion on the optimization method we've used and we present a comparison of our segmentation method with two other
  related works, and illustrate the results with several examples. Moreover, we have included a new local combinatorial
  optimization approach for elastica minimization. Finnaly,  the code produced in this paper is freely available on
  github\footnote{https://www.github.com/danoan/BTools}.}

\subsection{Outline}.
Section two reviews the concept of multigrid convergence and highlights its importance for the
definition of digital estimators. Next, we describe two convergent estimators used in this paper, one for tangent and
the other for curvature. They are used in the optimization model and in the definition of the digital elastica. Section
three describes a local combinatorial optimization model suitable for different curvature estimators. Section four
describes the proposed curvature evolution model along with several illustrations of digital flows. Section five
explains how to use the evolution model as a post-processing step in an image segmentation framework.

@@

%% for reference

%\begin{itemize}
%% \item{Geometric properties and regularization}
%% \item{The role of curvature}

%\item{\textbf{Constraint optimization}}
 % \begin{itemize}
  %% \item{Curvature regularity for region-based image segmentation and inpainting: A linear programming relaxation \cite{schoenemann09linear} }	
  %% \item{A Fast Algorithm for Euler's Elastica Model Using Augmented Lagrangian Method \cite{tai11elastica} }		
  %%\item{The elastic ratio: Introducing curvature into ratio-based image segmentation \cite{schoenemann2011elastic}
  %%}		%% DONE
 % \end{itemize}

%%\item{\textbf{Discrete approaches}}
%%  \begin{itemize}
  %% \item{Fast global optimization of curvature \cite{zehiry10fast}	} %% DONE
  %%\item{Efficient Squared Curvature \cite{nieuwenhuis14efficient} 	}
  %%\item{Contrast driven {E}lastica for image segmentation \cite{el2016contrast} }
  %%\item{Curvature Regularization for Curves and Surfaces in a Global Optimization Framework \cite{strandmark11globalframework} }

 % \end{itemize}

%%\end{itemize}
